---
title: "Computer Lab 1"
author: "Rachel Wright"
date: "1/30/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The commands we'll use today are part of the `tidyverse` package. This package should already be installed on the classroom computers.

If you try to run this code on your own computer, install tidyverse using the following command. Remove the "#" to make the code executable.

```{r install tidyverse}
#install.packages("tidyverse")
```

Load the `tidyverse` package. You only have to INSTALL a package once per computer, but you have to load it every time you want to use it.

```{r load tidyverse}
library(tidyverse)
```

# PART 1: INTRODUCTION TO R

You can use R like a calculator.

```{r calculator}
1+2
```

Use the next "chunk" to perform your own simple mathematical operations in R.

```{r student calculator}
# Do some math here


```

You can save the results to a "variable" (a named object that stores data for you).

```{r creating variables}
mySum <- 1+2 
mySum
```

Use the next couple of lines to save the output of a mathematical operation to a variable of your choice. (Hint: your variable name shouldn't contain spaces or special characters. "_" and "." are OK and are often used in place of a space.)

```{r student variable creation}
# Create a variable and do some math here


```

You can manipulate variables

```{r manipulate variables}
mySum*2
```

You can use "functions" embedded in the program to perform tasks.

```{r use function}
mySum <- sum(1,3)
mySum # note that I just over-wrote the old value of "mySum" by naming it the same name! 
# Be careful!!
```


You an create "vectors" of values.

```{r create vector}
myNumbers <- c(1,2,5)
```

And you can perform the same operations to vectors that you did to single values.

```{r use function on vector}
myNumbers*3
sum(myNumbers)
```

It's important to know what "type" of data you are working with in R. We already saw a VECTOR.

```{r vector data type}
myVector <- c(1:10)
myVector
```

You can also have a DATA FRAME.

```{r data frame data type}
myDataFrame <- data.frame("Category_Number" = c(1:4), 
                          "Category_Color" = c("red", "green", "yellow", "pink"))
myDataFrame
```

Notice that the data frame can contain different types of data.

"Category Number" contains integers (whole numbers).

"Category Color" contains factors (non-numeric values).

```{r stratify data frame}
str(myDataFrame)
sum(myDataFrame$Category_Number)
sum(myDataFrame$Category_Color)
```

On the other hand, a MATRIX has the same type of data in every cell.

```{r matrix}
myMatrix <- matrix(1:24, ncol=3)
myMatrix
```

# PART 2A: HISTOGRAM OF PHENOTYPIC VALUES AT A DIALLELIC LOCUS
This is a big jump in R skill... I know! Don't worry if it doesn't make sense right now.

Set the genotypes at a single diallelic locus.

For our purposes, this genotype is also the "phenotypic value".

```{r create genotypes for one locus}
genotypes <- data.frame("genotype" = c(0, 1, 1, 2)) 
# I had to make `genotypes` a data frame with only one column of data.
# This is for annoying reasons you don't need to worry about :)
genotypes
```

Plot the histogram for a single locus.

```{r plot values for single locus}
ggplot(genotypes, aes(x = genotype)) + 
  geom_histogram(aes(y = stat(count/sum(count)))) + 
  labs(title = "Phenotypic Values for a Single Diallelic Locus", 
       y = "Proportion", x = "Phenotypic Value") +
  theme_bw()
```

# PART 2B: HISTOGRAM OF PHENOTYPIC VALUES AT MULTIPLE DIALLELIC LOCI

Set the genotypes at a single diallelic locus.

For our purposes, this genotype is also the "phenotypic value."

```{r create genotypes for each locus}
genotypes <- c(0,1,1,2)
```

Set the possible genotypes for each locus.

Add as many loci as you want!

```{r assign genotypes to each locus}
locus_1 <- genotypes
locus_2 <- c(0,10,10,10)
locus_3 <- genotypes
locus_4 <- genotypes
locus_5 <- genotypes
locus_6 <- genotypes
locus_7 <- genotypes
```

Determine the number of possible combinations for every locus.

Don't forget to add your loci here!

```{r determine all combinations of genotypes}
all_combinations <- expand_grid(locus_1,locus_2,locus_3)
all_combinations
```

Make a histogram of phenotypic values.

```{r make a histogram}
all_combinations%>% mutate(phenotypic_value = rowSums(.)) %>% 
  ggplot(aes(x = phenotypic_value)) + 
  geom_histogram(aes(y = stat(count/sum(count)))) + 
  labs(title = "Phenotypic Values for Multiple Diallelic Loci", 
       y = "Proportion", x = "Phenotypic Value") +
  theme_bw()
```


Calculate the mean and standard deviation.

```{r calculate summary stats}
summary_stats <- all_combinations %>% mutate(phenotypic_value = rowSums(.)) %>% 
  summarize(mean = mean(phenotypic_value), sd = sd(phenotypic_value))
```

Plot a normal distribution.

```{r plot normal distribution}
all_combinations %>% mutate(phenotypic_value = rowSums(.)) %>% 
  ggplot(aes(x = phenotypic_value)) + 
  geom_histogram(aes(y = stat(count/sum(count)))) + 
  labs(title = "Phenotypic Values for Multiple Diallelic Loci", 
       y = "Proportion", x = "Phenotypic Value") +
  theme_bw() +
  stat_function(
    fun = dnorm, args = list(mean = summary_stats$mean, sd = summary_stats$sd),
    col = 'red')
```

